Modeling the Evolution of Mimicry

### 2016-02-26 docs/paper/Paper.tex
Updating predator table

\begin{table}[h]
\centering
\begin{tabular}{| p{1.5cm} | >{\centering} p{1.6cm} | p{4cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		ISize & 6 & Number of cell in single dimension of the 3-D toroidal cube\\ \hline
		World Size & 20 & Size of a single dimension of the 3-D toroidal cube\\ \hline
		Cell Size & \( World Size\) / \(ISize \) & Size of each cell\\ \hline
		Total Number of Cells & \( ISize^3  = 216\) & Total number of cells in the environment\\ 
	\hline
\end{tabular}
\caption{Parameters to control the environment.}
\label{tab:environment-control-parameters}
\end{table}

@@

Only the significant characteristics which are required for a successful model of mimicry is implemented. As for this simulation we are considering heliconius butterfly, the representation of their wing pattern is with the help of cellular automata. So basically every prey organism will contain a genetic representation of cellular automata with which the predator will identify the prey and store its level of palatability in memory.

%More content needs to be added for cellular automata.
\paragraph{Cellular Automata}

Cellular automata are computer generated patterns applying very simple rules. First invented by Jon von Neuman and later extended by Stephen Wolfram, this field of study has created a lot of enthusiasm in the evolution of complexity with blocks of simplicity, in this case rules. We choose cellular automata as it can be easily represented with the help of a binary genome and then evolutionary operations on the genomic representation, such as mutation and crossover can easily be applied.

@@

With CA based pattern representation, population of prey species with a specific CA pattern can be grouped as one single species. And by restricting inter species reproduction we can control the diversity of patterns. But there is mutation applied when similar species mate with each other, so new species do born out of generations of existing species. So that is why we have two separate mutation rate while reproducing prey species. One being the ``Pattern Mutation Rate" (default values are mentioned in table \ref{tab:prey-control-parameters}) with which we control mutation of the first 8 bits of the genome while the ``Genome Mutation Rate" is used to control mutation of rest of the 9 bit genome (table \ref{tab:prey-genome}). Similar efforts of multiple mutation rate at varying location has been used in developing Echo \citep{hraber1997}.

@@

\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{|c|c|c|c|}
	\hline
		\textbf{Pattern(8)} & \textbf{Palatability(2)} & \textbf{Mobility(6)} & \textbf{Reproduction(1)} \\ \hline
		10101101					 	& 							01		 		 & 			110001					&					1						 		 \\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 17 bit prey genome.}
\label{tab:prey-genome}
\end{table}

@@

\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|}
		\hline
			\textbf{Gene (Index 8 to 9)} &	\textbf{Palatable} \\ \hline
			00									& True 			\\ \hline
			01									& True 			\\ \hline
			10									& False 		\\ \hline
			11									& False 		\\
		\hline
	\end{tabular}
	\caption{Genetic representation of palatability}
	\label{tab:genetic-representation-palatability}
\end{table}

@@

\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Prey Size & 2 to 5 & Size of the prey species in the 3D FormAL  environment.\\ \hline
		Reproduction age limit & 100 & Minimum number of iterations or time steps a prey species need to be present in the simulation to get reproduction capability\\ \hline
		Reproduction interval & 1000 & Number of iterations a prey need to wait before reproducing again.\\ \hline
		Pattern Mutation Rate & 0.05 & Rate of Mutation of the pattern genome.\\ \hline
		Genome Mutation Rate & 0.5 & Rate of mutation of the rest of the genome excluding the pattern gene.\\ \hline
		Demise Age & 2000 & Age at which the prey species will be removed from the environment.\\
	\hline
\end{tabular}
\caption{Parameters to control prey population and visibility.}
\label{tab:prey-control-parameters}
\end{table}

@@

Predators in the system are designed to provide selection pressure to \textit{models} and \textit{mimics} for the evolution of mimicry. Similar to prey species, they are agents in the FormAL environment capable of mobility and reproduction. In addition to it, these agents are equipped with Hopfield Network Memory to be able to learn and recognize patterns of the prey species. Their mobility and reproduction capability are controlled at the genetic level, while their memory is not genetically controlled, as it was not possible to come up with a genetic representation for Hopfield Network. So every new predator species gets to born with zero memory and with no inheritance from parents. A set of parameters are defined to control predators' population and learning ability in the environment. Details of these parameters are given in table \ref{tab:predator-control-parameters}.

@@

\paragraph{Predator attack algorithm}
As soon as a predator reaches its attack age it selects random prey species around its vicinity and starts attacking them. This attack process also involves recognition of prey pattern, which is at the heart of mimicry simulation for this research. Two parameters have been defined to limit predator memorization and recognition process as both of these processes are computationally expensive. The ``Hopfield Minimum Memory Size" (default value in table \ref{tab:predator-control-parameters}) is the number of memory a predator needs to store before making intelligent decisions about attacking a new prey species. After a predator is born, it start attacking prey without any caution. But after every attack the predator will store its pattern and palatability level inside its memory. As soon as it reaches the minimum memory size, it will start making intelligent decision about attacking the next species. It will try to recognize the pattern and if found palatable, prey will be consumed. Otherwise prey will be thrown back into the environment. If the pattern is not recognized predator will try to consume it and in the process will store its palatability and pattern into memory.

In this way the predator memory is limited to \textsl{``Hopfield Maximum Memory Size"}. After reaching this memory predator will not store any more new pattern but will try to associate with the existing ones it has already stored.

@@

\begin{table}[h]
\centering
\begin{tabular}{|c|c|}
	\hline
		\textbf{Mobility(4)} & \textbf{Reproduction(1)} \\ \hline
				 1101					   &					1						 		\\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 5 bit predator genome.}
\label{tab:predator-genome}
\end{table}

@@

When the above two conditions are met, meaning the predator reaches its age for reproduction and also crosses its age interval, it randomly select another predator residing in the same cell. If this random predator is capable to reproduce depending on its 5th gene, then with a random single point crossover and mutation a new predator species is born which also resides on the same cell and initializes with zero memory configuration. Similarly when the new born predator reaches its maturity of reproductive age and if it is capable of reproduction then the process iterates itself.

@@

\subsection{Mimicry Ring Reports}
The mimicry ring reports consist entirely of the population of prey species categorized according to pattern and palatability. Data is stored at every instance of time of the simulation keeping an interval of 10 iterations. As the number of rings that get generated reaches as many as 50 or more, and all the population of ring do not last for the entire simulation, so while storing data we have taken the most populous of the surviving 8 rings to plot. Parameters are mentioned in table \ref{tab:ring-report-control-parameters}.

\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | >{\centering} p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Mimicry Ring Hamming Distance & 10 \% of the Pattern Size & If the Hamming distance between the pattern of the model and the mimic is 10 \% the size of the pattern then they are considered within the same ring.\\ \hline
		Number of Rings to report & 8 & This value is the number of most populous rings that are included in the report.\\
	\hline
\end{tabular}
\caption{Parameters to mimicry ring report.}
\label{tab:ring-report-control-parameters}
\end{table}

\subsection{Initial configuration with two prey species}
\label{subsec:init-conf-2prey}
% Put the table here
\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | p{1.5cm} | p{1cm} | p{.5cm} | p{1.5cm} | p{.5cm} |}
  \hline
   														&\multicolumn{3}{c|}{Prey configuration} 																	
   														& \multicolumn{2}{c|}{\parbox{2cm}{Predator \\ configuration}} \\ \hline
  \multirow{2}{*}{Population} & Rule110 (Palatable) & \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule110}} & 108 
  														& \multicolumn{2}{c|}{\multirow{2}{*}{10}} \\ \cline{2-4}
  					 									& Rule30 (Unpalatable)& \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule30}}  & 108 
  					 									& \multicolumn{2}{c|}{}\\ \hline
  \multirow{2}{*}{Reproduction} & Age Limit & \multicolumn{2}{c|}{100}  & \multicolumn{2}{c|}{500} \\ \cline{2-6}
  						 									& Interval  & \multicolumn{2}{c|}{1000} & \multicolumn{2}{c|}{1200} \\ \hline
  \multirow{2}{*}{Mutation Rate} & Pattern   & \multicolumn{2}{c|}{0.05} & \multicolumn{2}{c|}{\multirow{2}{*}{0.3}} \\ \cline{2-4}
  						 									 & Genome    & \multicolumn{2}{c|}{0.5}  & \multicolumn{2}{c|}{} \\ \hline
  Demise Age	 									 & \multicolumn{3}{c|}{2000}							& \multicolumn{2}{c|}{2500} \\ \hline
  Minimum Attack Age						 & \multicolumn{3}{c|}{} 						    & \multicolumn{2}{c|}{500} \\ \hline
  \multirow{2}{*}{\parbox{2cm}{Memory Configuration}} & \multicolumn{3}{c|}{} 					& Minimum & 2 \\ \cline{5-6}
   																			& \multicolumn{3}{c|}{} 					& Maximum & 10 \\ \hline  
\end{tabular}
\caption{Agent configuration of 2 prey species}
\label{tab:config-table-2-prey}
\end{table}

@@

The memory configuration size for predators is a very interesting parameter. The minimum memory size is directly associated with the number of prey species with which we initiate the simulation. Otherwise evolution of mimicry is not observed. As mentioned in table \ref{tab:predator-control-parameters} the minimum memory size is the number of prey species predator would consume before starting to make decisive consumption of prey species. After the initial birth of a predator and when the minimum attack age is crossed, it starts consuming prey species present in its vicinity without making any judgment. At this point its memory size is zero. As long as the minimum memory size is not reached predators will blindly consume prey species and insert their CA pattern and associated palatability into its Hopfield memory bank. No later than the minimum memory size has been reached, predators will start making intelligent decision about consuming its prey species. When catching a prey if its memory tells that it is palatable, it will consume that prey. Otherwise if memory recognizes it to be unpalatable predator will certainly let it go.

Now as we know the behavior of Hopfield memory, a recognition result will always be achieved depending on the similarity of the patterns stored in memory. So when minimum memory size has been reached the predator will always make a decision based on the similarity of the prey pattern currently captured and the patterns stored in memory.

% Put the image
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime10k-2Prey}
	\caption[Population distribution of mimicry rings (2 prey species, 10k iterations)]{Population distribution of mimicry rings, initialized with 2 prey species, 10k iterations}
	\label{fig:plot-2-prey}
\end{figure}

The plot in Figure \ref{fig:plot-2-prey} is simulation time verses prey population after running it for 10000 iterations. With the initial configuration in the above table we can observe that multiple rings of prey population have been created. Two prey species are considered to be in a ring if their CA pattern have a Hamming distance within 10 bits. This parameter is mentioned in table \ref{tab:ring-report-control-parameters}. Population of palatable species has been represented with line curve while population of unpalatable species has been presented with dotted curve. Different signs of squares, triangles and diamonds have been used to distinguish between species of prey population. The simulation was initiated with two prey species having CA rule of 30 and 110 and being palatable and unpalatable consecutively.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime5k-2Prey}
	\caption[Population distribution of mimicry rings (2 prey species, 5k iterations)]{Population distribution of mimicry rings, initialized with 2 prey species, 5k iterations.}
	\label{fig:plot-2-prey-5k}
\end{figure}

The plot of figure \ref{fig:plot-2-prey-5k} is from the same experiment as of figure \ref{fig:plot-2-prey}, but only up to 5000 iterations. By closely observing the graph of figure \ref{fig:plot-2-prey-5k}, we could see the population of two species have started dropping at around time 500 when the initial predator population reaches the maturity for consuming prey species (effect of the parameter `Minimum Attack Age' from table \ref{tab:config-table-2-prey}). At around time 1000 the prey population starts reproducing as the population increases (controlled by `Reproduction Age Limit' from table \ref{tab:config-table-2-prey}). At the same time different other species of prey gets to be born with mutated CA patterns. There is a straight drop of prey population at time 2000, when all the prey species with which the simulation was initialized come to its demise age (controlled by `Prey Demise Age' at table \ref{tab:config-table-2-prey}). Over time the population of CA Rule 30 dominates the population (figure \ref{fig:plot-2-prey}) as most predators recognize it as unpalatable. Similarly a palatable population of CA Rule 30 or within the same ring of palatable species starts rising, while at one point overlaps the population of CA Rule 110 (Time: 4000 approx.). CA Rule 110 was initialized as a set of palatable species.

@@

%Put the number of rings picture:
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/ringSize10k-2Prey}
	\caption[Number of mimicry rings (2 prey species)]{Number of mimicry rings, initialized with 2 prey species.}
	\label{fig:ringSize10k-2Prey}
\end{figure}

@@

%CARule table
\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{|l|c|c|c|}
  \hline
  CA Rule & \(60 \equiv 00111100\) & \(61 \equiv 00111101\) & \(62 \equiv 00111110 \) \\ \hline
  Pattern & \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule60}} 
  				& \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule61}} 
  				& \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule62}}\\
  \hline
\end{tabular}
\caption{Difference in prey pattern genotype and phenotype}
\label{tab:diff-in-pattern}
\end{table}

@@

The screen shot in figure \ref{fig:screenshot-simTime7600-2-prey} is for one instance of time of the simulation. The red agents are predators while the black agents with different textured patterns are the prey species. According to their behavior most of the prey species are flocking together in a group while also being chased by the predators, whose sole purpose is to consume prey species.

%Screenshot of the simulation
\begin{figure}[h]
	\centering
	\label{fig:screenshot-simTime7600-2-prey}
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime7600}
	\caption[Graphical representation of the model (simulation time: 7600)]{Graphical representation of the model, simulation time: 7600.}
\end{figure}

@@

To evaluate the simulation at a more complex level we increase the prey population to 900, consisting of 6 different species with very different pattern configuration. To boost predator-prey interaction we also increase the number of predator population to 30. Predator reproduction interval has been set comparatively lower in this simulation to 2000 iterations. As the initial population of prey species is very high, over time we are expecting even larger number of prey species. If predator population does not increase at similar rate there might be a prey population explosion were predator will have no effect on providing selection pressure for the evolution of mimicry. So that is why predator population control parameter has been set to this level. For memory configuration, the minimum number is set to 6, meaning the predator will have to store 6 different prey pattern configuration before starting to make intelligent decision about consuming them. As previously mentioned this number is always set in accordance with the initial number of different prey configuration.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime10k-6Prey}
	\caption[Population distribution of mimicry rings (6 prey species)]{Population distribution of mimicry rings, initialized with 6 prey species.}
	\label{fig:plot-6-prey}
\end{figure}

With the above set of parameters we receive the plot of figure \ref{fig:plot-6-prey}. An enormous diversity of species can be observed from this experiment. In addition to the six prey species with which the simulation starts, the total number of mimicry rings reach nearly 50. Observing the top 8 as presented in the plot of figure \ref{fig:plot-6-prey}, effects of Batesian mimicry is obvious. Unpalatable CA Rule 190 and 30 dominates the population. Similarly unpalatable CA Rule 105 is also dominant. But after 10000 iterations CA Rule 55 seems to have higher population than 105 even though 55 is a \textsl{palatable} set of species. This anomaly could be the reason for storing large number of patterns in the predator Hopfiled memory. As we know that the number of patterns stored in Hopfiled network is inversely proportional to its recognition rate. In this simulation memory size is controlled by the `Maximum Memory Configuration' which is set to 10. Also the value of `Minimum Memory Configuration' is 6, which is rather a large number. So there could be possibility of wrong recognition by Hopfield memory. Then again this phenomenon is common if we consider behavior of predators in nature. It is definitely hard for a single predator to memorize multiple poisonous patterns and because of which Mullerian mimicry takes effect.

@@

To further observe the effects of mimicry ring we initialize the simulation with all unpalatable prey species. After finding different anomalous behavior of predator recognition capability in the previous section for increased number of diverse prey species; for this experiment the initial number of species have been reduced from six to four. As explained earlier the minimum memory configuration is also set to four in accordance to the initial number of prey species. Rest of the parameters remain quite unchanged.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime8k-4Prey-unp}
	\caption[Population distribution of mimicry rings(4 prey species all unpalatable)]{Population distribution of mimicry rings, initialized with 4 prey species all unpalatable.}
	\label{fig:plot-4-prey-unp}
\end{figure}

@@

This experiment is an ideal scenario for observing Mullerian mimicry. As mentioned in section \ref{subsec:mullerian-mimicry}, Mullerian mimicry occurs between multiple species of unpalatable prey population. From the second work of Franks and Noble \citep{franks2003} mentioned in section \ref{subsubsec:models-by-frank-and-noble}, we note that multiple Mullerian mimicry rings are expected to converge into one large ring through the evolutionary process of punctuated equilibrium. But in this experiment as the predator's `Minimum Memory Configuration' is set to four, all predators have the capability to recognize four prey patterns before starting to make intelligent decision of consuming them. By setting `Minimum Memory Configuration' to one, also increasing `Predator Demise Age' to 7000 and decreasing predator's `Reproduction Age Interval' to 1500, we get the following plot.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime6k-4Prey-unp-1-mem}
	\caption[Population distribution of mimicry rings(4 prey, all unpalatable but reduced predator memory)]{Population distribution of mimicry rings, initialized with 4 prey species all unpalatable. Minimum Memory size reduced to one.}
	\label{fig:plot-4-prey-unp-1-mem}
\end{figure}

Running the simulation for 6000 iterations there was no sign for all prey population to converge into one large ring. As it can be observed in figure \ref{fig:plot-4-prey-unp-1-mem} all four unpalatable prey population have a very dominant presence in the simulation. Even though predator minimum memory has been reduced to only one pattern, different population of predators become familiar with different prey patterns, which results in the existence of multiple Mullerian mimicry ring instead of a single one.

\subsection{Initial configuration with only palatable species}
\label{subsec:init-only-palatable-species}

The parameters for this simulation has been set exactly the same as the previous section, except all the species are palatable at this point. All predator configuration has also been set to exactly same as before.

The results can be observed in figure \ref{fig:plot-4-prey-p} where all population of prey species have reached its demise at nearly 7000 iterations. By analyzing details we can see, around simulation time 500 all population of species come to a steady downfall as at this point initial predator species reach its' age for consumption. Around simulation time 1000 all population of prey species start increasing steadily because they reach their maturity level of reproduction. At iteration time 2000 there is a sudden drop of prey population because the initial population of prey have reached its demise age at this point. By this time the population of unpalatable species mutated from the palatable ones start increasing, and slowly around time 7000 and onwards the entire population of prey come to its demise. Predator population, from the beginning of the simulation has taken a very dominant effect. As all the species are palatable, predator's consumption and reproduction rate is extremely high and it keep increasing in a geometric rate.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/simTime10k-4Prey-p}
	\caption[Population distribution of mimicry rings (4 prey species all palatable)]{Population distribution of mimicry rings, initialized with 4 prey species all palatable.}
	\label{fig:plot-4-prey-p}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/ringSize10k-4Prey-p}
	\caption[Number of mimicry rings (4 prey species all palatable)]{Number of mimicry rings, initialized with 4 prey species all palatable.}
	\label{fig:ringSize8k-4-Prey-p}
\end{figure}

Observation of mimicry rings also tell us about the results we have seen in the plot of figure \ref{fig:plot-4-prey-p}. The number of rings keep increasing but comes to a sudden drop around simulation time 7000 and goes to zero.

>>>>

\begin{table}[h]
\small
\centering
\begin{tabular}{| p{1.5cm} | >{\centering} p{1.6cm} | p{4cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		ISize & 6 & Number of cell in single dimension of the 3-D toroidal cube\\ \hline
		World Size & 20 & Size of a single dimension of the 3-D toroidal cube\\ \hline
		Cell Size & \( World Size\) / \(ISize \) & Size of each cell\\ \hline
		Total Number of Cells & \( ISize^3  = 216\) & Total number of cells in the environment\\ 
	\hline
\end{tabular}
\caption{Parameters to control the environment.}
\label{tab:environment-control-parameters}
\end{table}

@@

Only the significant characteristics which are required for a successful model of mimicry is implemented. As for this simulation we are considering heliconius butterfly, the representation of their wing pattern is with the help of cellular automata. So every prey organism will contain a genetic representation of cellular automata with which the predator will identify the prey and store its level of palatability in memory. We choose cellular automata as it can be easily represented with the help of a binary genome and then evolutionary operations on the genomic representation, such as mutation and crossover can easily be applied.

@@

With CA based pattern representation, population of prey species with a specific CA pattern can be grouped as one single species. And by restricting inter species reproduction we can control the diversity of patterns. But there is mutation applied when similar species mate with each other, so new species do born out of generations of existing species. That is why we have two separate mutation rate while reproducing prey species. One being the ``Pattern Mutation Rate" (default values are mentioned in table \ref{tab:prey-control-parameters}) with which we control mutation of the first 8 bits of the genome while the ``Genome Mutation Rate" is used to control mutation of rest of the 9 bit genome (table \ref{tab:prey-genome}). Similar efforts of multiple mutation rate at varying location has been used in developing Echo \citep{hraber1997}.

@@

\begin{table}[h]
\small
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{|c|c|c|c|}
	\hline
		\textbf{Pattern(8)} & \textbf{Palatability(2)} & \textbf{Mobility(6)} & \textbf{Reproduction(1)} \\ \hline
		10101101					 	& 							01		 		 & 			110001					&					1						 		 \\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 17 bit prey genome.}
\label{tab:prey-genome}
\end{table}

@@

\begin{table}[h]
	\small
	\centering
	\begin{tabular}{|c|c|}
		\hline
			\textbf{Gene (Index 8 to 9)} &	\textbf{Palatable} \\ \hline
			00 or 01						& True 			\\ \hline
			10 or 11						& False 		\\
		\hline
	\end{tabular}
	\caption{Genetic representation of palatability}
	\label{tab:genetic-representation-palatability}
\end{table}

@@

\begin{table}[h]
\small
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Prey Size & 2 to 5 & Size of the prey species in the 3D FormAL  environment.\\ \hline
		Reproduction age limit & 100 & Minimum number of iterations or time steps a prey species need to be present in the simulation to get reproduction capability\\ \hline
		Reproduction interval & 1000 & Number of iterations a prey need to wait before reproducing again.\\ \hline
		Pattern Mutation Rate & 0.05 & Rate of Mutation of the pattern genome.\\ \hline
		Genome Mutation Rate & 0.5 & Rate of mutation of the rest of the genome excluding the pattern gene.\\ \hline
		Demise Age & 2000 & Age at which the prey species will be removed from the environment.\\
	\hline
\end{tabular}
\caption{Parameters to control prey population and visibility.}
\label{tab:prey-control-parameters}
\end{table}

@@

Predators in the system are designed to provide selection pressure to \textit{models} and \textit{mimics} for the evolution of mimicry. Similar to prey species, they are agents in the FormAL environment capable of mobility and reproduction. In addition to it, these agents are equipped with Hopfield Network Memory to be able to learn and recognize patterns of the prey species. Their mobility and reproduction capability are controlled at the genetic level, while their memory is not genetically controlled, as it was not possible to come up with a genetic representation for Hopfield Network. So every new predator species gets to born with zero memory and with no inheritance from parents. A set of parameters are defined to control predators' population and learning ability in the environment. Details of which are in table \ref{tab:predator-control-parameters}.

@@

\paragraph{Predator attack algorithm}
As soon as a predator reaches its attack age it selects random prey species around its vicinity and starts attacking them. This attack process also involves recognition of prey pattern. Two parameters have been defined to limit predator memorization and recognition process as both of these processes are computationally expensive. The ``Hopfield Minimum Memory Size" is the number of memory a predator needs to store before making intelligent decisions about attacking a new prey species. After a predator is born, it start attacking prey without any caution. But after every attack the predator will store its pattern and palatability level inside its memory. As soon as it reaches the minimum memory size, it will start making intelligent decision about attacking the next species. It will try to recognize the pattern and if found palatable, prey will be consumed. Otherwise prey will be thrown back into the environment. If the pattern is not recognized predator will try to consume it and in the process will store its palatability and pattern into memory. In this way the predator memory is limited to \textsl{``Hopfield Maximum Memory Size"}. After reaching this memory predator will not store any more new pattern but will try to associate with the existing ones it has already stored.

@@

\begin{table}[h]
\small
\centering
\begin{tabular}{|c|c|}
	\hline
		\textbf{Mobility(4)} & \textbf{Reproduction(1)} \\ \hline
				 1101					   &					1						 		\\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 5 bit predator genome.}
\label{tab:predator-genome}
\end{table}

@@

@@

\subsection{Mimicry Ring Reports}
The mimicry ring reports consist entirely of the population of prey species categorized according to pattern and palatability. Data is stored for time interval of 10 iterations. As the number of rings that get generated reaches as many as 50 or more, and all the population of ring do not last for the entire simulation, so while storing data we have taken the most populous of the surviving 8 rings to plot. Parameters are mentioned in table \ref{tab:ring-report-control-parameters}.

\begin{table}[h]
\small
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | >{\centering} p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Mimicry Ring Hamming Distance & 10 \% of the Pattern Size & If the Hamming distance between the pattern of the model and the mimic is 10 \% the size of the pattern then they are considered within the same ring.\\ \hline
		Number of Rings to report & 8 & This value is the number of most populous rings that are included in the report.\\
	\hline
\end{tabular}
\caption{Parameters to mimicry ring report.}
\label{tab:ring-report-control-parameters}
\end{table}

\subsection{Initial configuration with two prey species}
\label{subsec:init-conf-2prey}
% Put the table here
\begin{table}[h]
\small
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | p{1.5cm} | p{1cm} | p{.5cm} | p{1.5cm} | p{.5cm} |}
  \hline
   														&\multicolumn{3}{c|}{Prey configuration} 																	
   														& \multicolumn{2}{c|}{\parbox{2cm}{Predator \\ configuration}} \\ \hline
  \multirow{2}{*}{Population} & Rule110 (Palatable) & \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule110}} & 108 
  														& \multicolumn{2}{c|}{\multirow{2}{*}{10}} \\ \cline{2-4}
  					 									& Rule30 (Unpalatable)& \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule30}}  & 108 
  					 									& \multicolumn{2}{c|}{}\\ \hline
  \multirow{2}{*}{Reproduction} & Age Limit & \multicolumn{2}{c|}{100}  & \multicolumn{2}{c|}{500} \\ \cline{2-6}
  						 									& Interval  & \multicolumn{2}{c|}{1000} & \multicolumn{2}{c|}{1200} \\ \hline
  \multirow{2}{*}{Mutation Rate} & Pattern   & \multicolumn{2}{c|}{0.05} & \multicolumn{2}{c|}{\multirow{2}{*}{0.3}} \\ \cline{2-4}
  						 									 & Genome    & \multicolumn{2}{c|}{0.5}  & \multicolumn{2}{c|}{} \\ \hline
  Demise Age	 									 & \multicolumn{3}{c|}{2000}							& \multicolumn{2}{c|}{2500} \\ \hline
  Minimum Attack Age						 & \multicolumn{3}{c|}{} 						    & \multicolumn{2}{c|}{500} \\ \hline
  \multirow{2}{*}{\parbox{2cm}{Memory Configuration}} & \multicolumn{3}{c|}{} 					& Minimum & 2 \\ \cline{5-6}
   																			& \multicolumn{3}{c|}{} 					& Maximum & 10 \\ \hline  
\end{tabular}
\caption{Agent configuration of 2 prey species}
\label{tab:config-table-2-prey}
\end{table}

@@

The memory configuration size for predators is a very interesting parameter. The minimum memory size is directly associated with the number of prey species with which we initiate the simulation. Otherwise evolution of mimicry is not observed. As mentioned in table \ref{tab:predator-control-parameters} the minimum memory size is the number of prey species predator would consume before starting to make decisive consumption of prey species. Based on the behavior of Hopfield memory, a recognition result will always be achieved depending on the similarity of the patterns stored in memory. So when minimum memory size has been reached the predator will always make a decision based on the similarity of the prey pattern currently captured and the patterns stored in memory.

% Put the image
\begin{figure*}[t!]
	\centering
	\includegraphics[width=0.7\textwidth]{../tex/images/simTime10k-2Prey}
	\caption[Population distribution of mimicry rings (2 prey species, 10k iterations)]{Population distribution of mimicry rings, initialized with 2 prey species, 10k iterations}
	\label{fig:plot-2-prey}
\end{figure*}

The plot in Figure \ref{fig:plot-2-prey} is simulation time verses prey population after running it for 10000 iterations. With the initial configuration in the above table we can observe that multiple rings of prey population have been created. Two prey species are considered to be in a ring if their CA pattern have a Hamming distance within 10 bits. Population of palatable species has been represented with line curve while population of unpalatable species has been presented with dotted curve. Different signs of squares, triangles and diamonds have been used to distinguish between species of prey population. The simulation was initiated with two prey species having CA rule of 30 and 110 and being palatable and unpalatable consecutively. Over time the population of CA Rule 30 dominates the population (Figure \ref{fig:plot-2-prey}) as most predators recognize it as unpalatable. Similarly a palatable population of CA Rule 30 or within the same ring of palatable species starts rising, while at one point overlaps the population of CA Rule 110 (Time: 4000 approx.). CA Rule 110 was initialized as a set of palatable species.

@@

%Put the number of rings picture:
\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{../tex/images/ringSize10k-2Prey}
	\caption[Number of mimicry rings (2 prey species)]{Number of mimicry rings, initialized with 2 prey species.}
	\label{fig:ringSize10k-2Prey}
\end{figure}

@@

%CARule table
\begin{table}[h]
\small
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{|l|c|c|c|}
  \hline
  CA Rule & \(60 \equiv 00111100\) & \(61 \equiv 00111101\) & \(62 \equiv 00111110 \) \\ \hline
  Pattern & \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule60}} 
  				& \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule61}} 
  				& \parbox[c]{2.1em}{\includegraphics[scale=0.50]{../tex/images/CARule62}}\\
  \hline
\end{tabular}
\caption{Difference in prey pattern genotype and phenotype}
\label{tab:diff-in-pattern}
\end{table}

@@

@@

To evaluate the simulation at a more complex level we increase the prey population to 900, consisting of 6 different species with very different pattern configuration. To boost predator-prey interaction we also increase the number of predator population to 30. Predator reproduction interval has been set comparatively lower in this simulation to 2000 iterations. As the initial population of prey species is very high, over time we are expecting even larger number of prey species. If predator population does not increase at similar rate there might be a prey population explosion were predator will have no effect on providing selection pressure for the evolution of mimicry. So that is why predator population control parameter has been set to this level. For memory configuration, the minimum number is set to 6, meaning the predator will have to store 6 different prey pattern configuration before starting to make intelligent decision about consuming them. As previously mentioned this number is always set in accordance with the initial number of different prey configuration. An enormous diversity of species can be observed from this experiment. In addition to the six prey species with which the simulation starts, the total number of mimicry rings reach nearly 50.

@@

To further observe the effects of mimicry ring we initialize the simulation with all four unpalatable prey species. As explained earlier the minimum memory configuration is also set to four in accordance to the initial number of prey species. Rest of the parameters remain quite unchanged.

\begin{figure*}[t]
	\centering
	\includegraphics[width=0.7\textwidth]{../tex/images/simTime8k-4Prey-unp}
	\caption[Population distribution of mimicry rings(4 prey species all unpalatable)]{Population distribution of mimicry rings, initialized with 4 prey species all unpalatable.}
	\label{fig:plot-4-prey-unp}
\end{figure*}

@@

This experiment is an ideal scenario for observing Mullerian mimicry. Mullerian mimicry occurs between multiple species of unpalatable prey population. From the second work of Franks and Noble \citep{franks2003} mentioned in section \ref{subsubsec:models-by-frank-and-noble}, we note that multiple Mullerian mimicry rings are expected to converge into one large ring through the evolutionary process of punctuated equilibrium. But in this experiment as the predator's `Minimum Memory Configuration' is set to four, all predators have the capability to recognize four prey patterns before starting to make intelligent decision of consuming them. By setting `Minimum Memory Configuration' to one, also increasing `Predator Demise Age' to 7000 and decreasing predator's `Reproduction Age Interval' to 1500, we run the simulation for 6000 iterations, and there was no sign for all prey population to converge into one large ring. All four unpalatable prey population have a very dominant presence in the simulation. Even though predator minimum memory has been reduced to only one pattern, different population of predators become familiar with different prey patterns, which results in the existence of multiple Mullerian mimicry ring instead of a single one.

In contrast when the simulation was initiated with only palatable species all population of prey have reached its demise at nearly 7000 iterations.
